Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.

### 2023-02-10 manuscript.Rmd
rephare statistical relevance

1. Researchers judge the evidence for or against a hypothesis rationally.
2. They expect other researchers to apply the same rational process.
2. All else being equal, researchers try to maximize the expected epistemic value for other researchers.

The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
Our rationale is as follows.
Researchers who decide to conduct a certain study are actually choosing a study to bet on.
They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in what studies to conduct.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
That is epistemic value is gained if evidence is observed, given that it is more likely to observe evidence when the theory holds than if the theory does not hold.

>>>>

1. Researchers judge the evidence for or against a hypothesis rationally.
2. They expect other researchers to apply a similar rational process.
2. They try to maximize the expected epistemic value for other researchers.

The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
Our rationale is as follows.
Researchers who decide to conduct a certain study are actually choosing a study to bet on.
They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in what studies to conduct.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
That is, evidence only increases epistemic value for a theory, if the evidence is more likely to observe under the theory than without it.
